Product plots

### 2011-06-23 reviews/response.txt

We are grateful for the helpful feedback provided by the reviewers. We have added a related work section, including many of the missing references we missed. We have also weakened our claims about the superiority of rectangles, instead providing a more pragmatic justification

Other changes:

* Rather than eliminating the section describing the R implementation altogether, we have expanded it slightly to give a taste of how it might be used.

------

### 2011-06-23 prodplots-infovis.tex
More on R package.  Start on response

The product plots framework is heavily influenced by the work of Wilkinson \citep{wilkinson:1999, wilkinson:2005}, who suggested that both mosaic plots and treemaps can be described as plots with coordinate systems based on recursive partitioning. Our work builds on this by adding conditioning, and defining a ``grammar of area plots'', a domain specific language for describing, in detail, this small family of plots. Other similar efforts to build domain specific languages for visualisation are APT \citep{mackinlay:1986} and VisQL \citep{mackinlay:2007}: our framework is driven by similar forces, but carves out a much smaller niche.

@@

We have provided a reference implementation of these ideas in an R \citep{R} package called {\tt productplots}, available from {\tt http://github.com/hadley/productplots}. There are two main functions: {\tt prodcalc}, which computes the coordinates of each rectangle; and {\tt prodplot}, which displays the rectangles with the ggplot2 package \citep{me:ggplot2}. The code is well tested, and ensures that the constraints are always satisfied. The {\tt productplots} code to used to create the images in this paper is available from the authors.

>>>>

The product plots framework is heavily influenced by the work of Wilkinson \citep{wilkinson:1999, wilkinson:2005}, who suggested that both mosaic plots and treemaps can be described as plots with coordinate systems based on recursive partitioning. Our work builds on this by adding conditioning, and defining a ``grammar of area plots'', a domain specific language for describing, in detail, this small family of plots. Other similar efforts to build domain specific languages for visualisation are APT \citep{mackinlay:1986} and VisQL \citep{mackinlay:2007}: our framework is driven by similar forces, but carves out a much smaller niche. This makes products plots less expressive, but more efficient, because the number of primitives required to describe the smaller domain is much smaller.

@@

We have provided a reference implementation of these ideas in an R \citep{R} package called {\tt productplots}, available from {\tt http://github.com/hadley/productplots}. There are two main functions: {\tt prodcalc}, which computes the coordinates of each rectangle; and {\tt prodplot}, which displays the rectangles with the ggplot2 package \citep{me:ggplot2}. The code is well tested, and ensures that the constraints are always satisfied.

For example, the following code conveys the essence of the plots in Figure~\ref{fig:part-cond}. The function {\tt prodplot} creates plots of the {\tt happy} data, defined using the standard R formula notation (\verb|~|), with the convention that {\tt |} denotes conditioning. The final argument lists the graphical primitives to use for display. This can also be a template function like {\tt mosaic()} which produces a standard, named, graphic.

\begin{verbatim}
prodplot(happy, ~ happy + finrela + health, 
  c("vspine", "fluct"))
prodplot(happy, ~ happy | finrela + health, 
  c("vspine", "fluct"))
prodplot(happy, ~ happy | finrela + health, 
  c("hspine", "fluct"))
\end{verbatim}

As for all presentation graphics, the actual code is somewhat more complicated, as we make a number of tweaks for optimal display. The complete {\tt productplots} code to used to create the images in this paper is available from the authors upon request.
